---
title: "p8105_hw6_jl5548"
author: "J L"
date: "November 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggridges)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```

## Problem 1

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
# load and clean variable names
df_birthweight = read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() 

# tidy data and store in another dataframe
df_birthweight_tidy = df_birthweight %>% 
  mutate(  # convert numeric to factor, and categorize to corresponding label
    babysex = factor(
      case_when(
        babysex == 1 ~ "male",
        babysex == 2 ~ "female"
      )
    ), 
    frace = factor(
      case_when(
        frace == 1 ~ "White",
        frace == 2 ~ "Black",
        frace == 3 ~ "Asian",
        frace == 4 ~ "Puerto Rican",
        frace == 8 ~ "Other",
        frace == 9 ~ "Unknown"
      )
    ),
    malform = factor(
      case_when(
        malform == 0 ~ "absent",
        malform == 1 ~ "present"
      )
    ), 
    mrace = factor(
      case_when(
        mrace == 1 ~ "White",
        mrace == 2 ~ "Black",
        mrace == 3 ~ "Asian",
        mrace == 4 ~ "Puerto Rican",
        mrace == 8 ~ "Other"
      )
    )
  )

# check for missing data
anyNA(df_birthweight)

# display tidied dataset
df_birthweight_tidy
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}

```



