---
title: "p8105_hw6_jl5548"
author: "J L"
date: "November 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
set.seed(1)
```

## Problem 1

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
# load and clean data
df_birthweight = read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(  # convert numeric to factor 
    babysex = factor(babysex), 
    frace = factor(frace),
    malform = factor(malform), 
    mrace = factor(mrace) 
  )
# check for missing data
anyNA(df_birthweight)
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}

```



